\section{Introduction}

\subsection*{Motivation}

Priority queues arise naturally wherever heterogeneous demand must be served by a shared resource: triage in emergency departments, differentiated service-level agreements in telecommunications networks, real-time versus best-effort traffic in operating system schedulers, and tiered customer support in call centres.  In each setting a key modelling question is how the steady-state queue lengths and waiting times depend on the priority discipline and on the parameters governing customer patience and mobility.

This thesis analyses a two-class non-preemptive priority $M/M/1$ queue---Poisson arrivals of rate $\lambda_i$ for class~$i$, exponential service of rate $\mu$ shared by both classes, and a single server that completes the current service before switching to a higher-priority customer---augmented with two \emph{departure} mechanisms that move or remove waiting customers before they reach the server.

\textbf{Jockeying.}  A waiting customer in one queue may switch to the other at a per-customer rate ($\gamma_1$ for the direction class-1$\to$class-2, $\gamma_2$ for the reverse direction).  Jockeying is a transfer: the total number of customers in the system is preserved, but the composition of each queue changes.  It models rational behaviour in systems with visible queue lengths, and has been studied in parallel-queue models \cite{adan2002queueing} and accumulating-priority systems \cite{stanford2014accumulating}.

\textbf{Abandonment (reneging).}  A waiting customer of class~$i$ holds an independent $\mathrm{Exp}(\theta_i)$ patience clock and leaves the system when it expires, at a per-customer aggregate rate $\theta_i n_i$.  Abandonment is a true departure: the total population decreases, the stability condition relaxes, and the class-2 queue is affected indirectly because the server is freed sooner.  It has been studied in multi-server priority models \cite{jouini2014multiple,deWaal1992approximate} and in scheduling contexts \cite{huchandong2022proactive,shmelev2025letting}.

The interplay of the two mechanisms within a priority discipline is the central object of this thesis.

\subsection*{The analytical challenge}

The combination of jockeying and abandonment in full generality defines Model~X, whose state space is $\{(0)\}\cup\{(n_1,n_2):n_1,n_2\ge0\}$ with the idle state $(0)$ and all states $(n_1,n_2)$ implying an active server.  The bivariate probability generating function $P(x,y)=\mathbb{E}[x^{N_1}y^{N_2}]$ (where $N_i$ counts waiting customers of class~$i$, \emph{excluding} the one in service) satisfies a functional equation that mixes three kinds of terms: an algebraic coefficient, a partial derivative in $x$ (from jockeying and abandonment in class~1), and a partial derivative in $y$ (from jockeying and abandonment in class~2).  When both derivatives are present and coupled, the equation belongs to a family of integro-differential equations for which no closed-form solution is known.

The strategy adopted here is to isolate and solve the \emph{tractable specializations} of Model~X, identified by zeroing all but one extra parameter:

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{clll}
\hline
\textbf{Model} & \textbf{Active mechanism} & \textbf{Equation type} & \textbf{Status} \\
\hline
A   & none                          & algebraic PGF equation    & solved (three routes) \\
B   & bidirectional jockeying        & PDE, Volterra family      & open (deliberate) \\
B$_2$ & one-way jockeying ($1\to2$) & ODE in $x$, Kummer        & solved \\
C$_2$ & class-1 abandonment         & ODE in $x$, Kummer        & solved \\
\hline
\end{tabular}
\end{center}

Model~B's incompleteness is structural, not accidental: with two-way jockeying present, the boundary condition on $P_y(y)$ reduces to a Volterra integral equation for which no closed-form kernel inversion is available.  Model~B$_2$---one-way jockeying only---removes the $\partial P/\partial y$ term and renders the functional equation an ODE, which is solvable.

\subsection*{Methods}

Three complementary proof strategies are employed throughout.

\textbf{Kernel method / ODE with integrating factor.}  For Model~A, the algebraic coefficient of $P(x,y)$ vanishes on a curve (the ``kernel''); evaluating the equation there determines $P_y(y)$ without solving a differential equation.  For Models~B$_2$ and~C$_2$, the reduction to a first-order linear ODE admits an explicit integrating factor.  The unknown boundary function $P_y(y)$ is then pinned by a regularity argument at a singularity of the integrating factor: an \emph{analyticity} (holomorphy) condition at the interior point $x=y$ for Model~B$_2$, and a \emph{boundedness} condition at the domain boundary $x=1$ for Model~C$_2$.

\textbf{Probabilistic argument via Pollaczek--Khinchine.}  For Models~A and~C$_2$, the Poisson character of the class-2 arrival stream is preserved by the extra mechanism: in Model~A because no mechanism is active, and in Model~C$_2$ because class-1 abandonment does not disturb the class-2 arrival process.  In both cases the PASTA property applies to class-2 customers, who then see an $M/G/1$ queue with effective service time equal to a busy period of the class-1 subsystem.  The Pollaczek--Khinchine formula delivers $P_y(y)$ directly, providing an independent confirmation of the analytic result.  Jockeying destroys this Poisson structure (customers now arrive at the class-2 queue from inside the system as well as from outside), so the probabilistic route is \emph{not available} for Models~B and~B$_2$.

\textbf{Partial PGF on the alternative state space $\widetilde{S}$.}  The system can also be parameterized by the pair $(N_2, N)$ where $N=N_1+N_2$ is the total waiting population.  This coordinatization, denoted $\widetilde{S}$, yields a different set of balance equations and a partial PGF $\widetilde{P}(y,n)$ that is a polynomial in $y$ for each fixed $n$.  The polynomial structure removes convergence issues and provides an alternative closed-form route for Model~A, presented in parallel for completeness.

\subsection*{Contributions}

The main results of this thesis are the following.

\begin{enumerate}
    \item \textbf{Three independent solutions of Model~A} (Section~\ref{sec:model_a}): a kernel-method proof, a probabilistic P-K proof, and a PPGF proof on $\widetilde{S}$.  Each route exploits a different structural property; all three converge to the same rational closed form.

    \item \textbf{Closed-form solution of Model~B$_2$} (Section~\ref{sec:model_b2}): $P(x,y)$ is expressed as an integral form whose integrand involves an integrating factor with a Kummer-function singularity at $x=y$.  The boundary function $P_y(y)$ is a ratio of confluent hypergeometric functions.  The diagonal $P(z,z)$ and the empty-state probabilities $\pi_0, \pi(0,0)$ coincide with those of Model~A for all $\gamma_1$: jockeying does not affect total-population statistics.

    \item \textbf{Closed-form solution of Model~C$_2$} (Section~\ref{sec:model_c2}): $P(x,y)$ is expressed in terms of two incomplete Beta integrals $\mathcal{I}_1, \mathcal{I}_2$ derived from the ODE solution, together with the LST $\widetilde{B}_C(s)$ of the busy period of the class-1 $M/M/1{+}M$ subsystem.  The boundary function $P_y(y)$ takes the Pollaczek--Khinchine form with $B_C$ as the effective service time.  The stability condition $\lambda_2\,\mathbb{E}[B_C]<1$ is strictly weaker than $\rho<1$ for all $\theta_1>0$.

    \item \textbf{Structural comparison} (Section~\ref{sec:comparison}): Models~B$_2$ and~C$_2$ share a common Erlang-A skeleton---both produce ${}_1F_1(1;\mu/r+1;\lambda_1/r)$ as a key parameter, with $r=\gamma_1$ or $r=\theta_1$---despite the different mechanisms by which $P_y(y)$ is pinned.  Jockeying and abandonment are shown to be structurally dual: one preserves total-population statistics but redistributes load between classes; the other reduces all queue lengths but admits no total-population conservation law.

    \item \textbf{Experiment models B$_2^1$ and C$_2^1$} (Section~\ref{sec:experiment}): variants in which only the head-of-line customer in Queue~1 is subject to the departure mechanism (at a fixed rate independent of queue length).  The resulting PGF equation is algebraic; the kernel method gives rational closed forms for $P(x,y)$, and these are validated against CTMC.
\end{enumerate}

All results are verified numerically against the exact stationary distribution of a truncated CTMC with $N_{\max}=40$, with maximum relative errors below $10^{-7}$ for every theorem.

\subsection*{Structure of the thesis}

Section~\ref{sec:literature} surveys the relevant literature.  Section~\ref{sec:prelim} collects the probabilistic and analytic background used throughout.  Section~\ref{sec:model_x} defines Model~X and derives the general balance equations and the fundamental PGF equation.  Sections~\ref{sec:model_a}--\ref{sec:model_c2} analyse Models~A, B, B$_2$, and C$_2$ in turn, each following the canonical template: stability, fundamental equation, theorem, and proof (analytic then probabilistic where applicable).  Section~\ref{sec:comparison} compares the three solved models.  Section~\ref{sec:results} presents the numerical validation and performance metrics.  Section~\ref{sec:experiment} introduces and solves the head-of-line variants.
