\section{Literature}\label{sec:literature}

% ── Overview ──────────────────────────────────────────────────────────────────
The study of two-class priority queues with departure mechanisms sits at the intersection of three well-developed streams: exact PGF analysis of non-preemptive priority queues, the theory of queues with customer impatience (abandonment), and the theory of jockeying or priority changes. This thesis contributes an exact, single-server treatment that combines jockeying and abandonment within a unified PGF framework, filling a gap that the multi-server and approximate literature has so far left open.

% ── §2.1 ──────────────────────────────────────────────────────────────────────
\subsection*{Non-preemptive priority queues: exact queue-length analysis}

The direct methodological ancestor of this work is \citet{cohen1956delay}, who introduced the partial PGF technique (PPGF) for a two-class priority queue loaded by two independent Poisson streams — the precursor of the Cohen trick used in \S\ref{ssec:mm1} and throughout Sections~5--7.

\citet{zuk2023explicit} carry this line of work to its most recent exact closed-form treatment: they derive the joint and marginal queue-length distributions for a non-preemptive M/M/$c$ queue with two priority levels (equal service rates) via PGF methods and a quadratic recurrence.
% [Similarities to this thesis:] non-preemptive priority, two classes, common
%  exponential service rate, PGF/Cohen methodology, queue-length distributions.
% [Differences:] $c \geq 1$ servers (thesis: single server only), no jockeying,
%  no abandonment; their single-server ($c=1$) case recovers Model~$A$ of this
%  thesis.
% [Cite in:] §5 (Model~$A$, after the theorem — note that the c=1 case coincides).

The preemptive multi-server case is analysed exactly by \citet{wang2015mmc}, who derive the queue-length PGF for an M/M/$c$ queue with two preemptive priority classes and class-specific service rates.
% [Similarities:] exact PGF analysis, two priority classes, Markov chain approach.
% [Differences:] preemptive discipline (this thesis: non-preemptive throughout),
%  multiple servers, different service rates per class, no abandonment/jockeying.
% [Cite in:] §2 Literature; §9 Comparison (contrast preemptive vs. non-preemptive).

% ── §2.2 ──────────────────────────────────────────────────────────────────────
\subsection*{Priority queues with abandonment}

The multi-server analog of Model~$C_2$ is studied by \citet{jouini2014multiple}, who consider an M/M/$s$+M queue with two non-preemptive priority classes and exponential patience times (Erlang-A structure per class). They derive Laplace- Stieltjes transforms for the unconditional and conditional waiting times, including the probability of service and abandonment.
% [Similarities:] non-preemptive priority, two classes, exponential abandonment
%  (Erlang-A), Markovian arrivals and service; their model is the multi-server
%  version of Model~$C$₂.
% [Differences:] multiple servers ($c \geq 1$); focus on waiting time LSTs
%  (this thesis: queue-length bivariate PGF); no jockeying; approximate methods
%  only for $c>1$.
% [Cite in:] §8 (Model~$C$₂ intro — position the single-server exact treatment
%  relative to the multi-server approximate one); §10 (Results).

\citet{shmelev2025letting} study a multi-server, multi-class queue where the operator can reject arriving customers or initiate (controlled) abandonment, in addition to customer-initiated impatience. Both Erlang-B (rejection) and Erlang-A (impatience) substructures appear, and a heavy-traffic index policy is shown to be asymptotically optimal.
% [Similarities:] Erlang-A (M/M/1+M) substructure per class, abandonment as
%  a departure mechanism, multi-class scheduling.
% [Differences:] controlled (operator-initiated) abandonment, multiple servers,
%  focus on scheduling policy and heavy-traffic asymptotics rather than exact
%  stationary distributions; no bivariate PGF.
% [Cite in:] §8 (Model~$C$₂ — contextualise Erlang-A substructure); §10.

% ── §2.3 ──────────────────────────────────────────────────────────────────────
\subsection*{Jockeying and priority changes}

\citet{deWaal1992approximate} study an M/G/1 priority queue where customers become impatient and \emph{change their priority class} over time. The analysis is approximate (exact solutions are intractable for general service times), providing LST-based waiting time approximations.
% [Similarities:] jockeying modelled as priority changes due to impatience;
%  non-preemptive priority; single server.
% [Differences:] approximate analysis (this thesis: exact); M/G/1 (this thesis:
%  M/M/1); priority change is class-upgrade over time, not queue-switch
%  (jockeying in this thesis moves a customer between two physical queues,
%  preserving total n); focus on waiting time, not queue-length PGF.
% [Cite in:] §6 (Model~$B$ preamble — motivate why exact jockeying is hard);
%  §2 Literature.

\citet{stanford2014accumulating} introduce the \emph{accumulating priority} queue, in which each class $i$ accumulates priority at rate $b_i$ per unit waiting time so that older customers overtake newer ones over time. They derive exact waiting time distributions for an M/G/1 queue via Laplace--Stieltjes transforms and a maximum-priority process.
% [Similarities:] dynamic priority changes (analogue of jockeying); single server;
%  non-preemptive discipline; exact analysis.
% [Differences:] priority \emph{accumulates continuously} (this thesis: discrete
%  jockeying events at rate γᵢ per customer); M/G/1 (this thesis: M/M/1);
%  focus on sojourn times, not queue-length PGF.
% [Cite in:] §6 (Model~$B$ — accumulating priority as a continuous-time analogue
%  of jockeying); §2 Literature.

\citet{huchandong2022proactive} model a multi-server queue where customers transition between a \emph{moderate} and an \emph{urgent} class while waiting, driven by deterioration and improvement dynamics. Optimal scheduling is characterised by a modified $c\mu/\theta$ index in heavy traffic.
% [Similarities:] dynamic class transitions while waiting (structurally similar
%  to jockeying); non-preemptive priority; impatience/abandonment modelled.
% [Differences:] class transitions are stochastic deterioration/improvement
%  (not customer-controlled queue-switching); multiple servers; focus on
%  optimal scheduling policy, not exact stationary distributions.
% [Cite in:] §6 (Model~$B$ preamble); §2 Literature.

% ── §2.4 ──────────────────────────────────────────────────────────────────────
\subsection*{Multi-server extensions and performance design}

\citet{baron2025rationing} study \emph{capacity rationing} in an M/M/$c$ non-preemptive priority queue: low-priority customers may only enter service when at least $k$ servers are idle. They show the optimal rationing level scales as $O(\sqrt{c})$ in the Halfin--Whitt regime.
% [Similarities:] M/M/c, non-preemptive priority, two classes, exact analysis
%  for heterogeneous service rates.
% [Differences:] multiple servers; capacity rationing mechanism (not jockeying
%  or abandonment); heavy-traffic regime; focus on policy design.
% [Cite in:] §9 Comparison / §2 Literature — as a different type of
%  inter-class interaction mechanism.

% ── §2.5 ──────────────────────────────────────────────────────────────────────
\subsection*{Position of this thesis}

The following table summarises the placement of this work relative to the
closest papers.

\begin{table}[H]
\centering
\caption{Literature overview: models and mechanisms.
\checkmark~= present, --~= absent, $\approx$~= approximate only.}
\label{tab:literature}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lccccccc}
\hline
\textbf{Reference} & \textbf{Servers} & \textbf{Non-preem.} & \textbf{Jockeying}
  & \textbf{Aband.} & \textbf{Queue-len.} & \textbf{Exact} \\
\hline
Cohen (1956)                      & 1   & \checkmark & -- & --         & \checkmark & \checkmark \\
de Waal (1992)                    & 1   & \checkmark & \checkmark & --  & --         & $\approx$ \\
Jouini \& Roubos (2014)           & $c$ & \checkmark & -- & \checkmark & --         & \checkmark \\
Wang et al.\ (2015)               & $c$ & --         & -- & --         & \checkmark & \checkmark \\
Stanford et al.\ (2014)           & 1   & \checkmark & \checkmark & --  & --         & \checkmark \\
Zuk \& Kirszenblat (2023)         & $c$ & \checkmark & -- & --         & \checkmark & \checkmark \\
Baron et al.\ (2025)              & $c$ & \checkmark & -- & --         & --         & \checkmark \\
Hu, Chan \& Dong (2022)           & $c$ & \checkmark & \checkmark & \checkmark & --   & $\approx$ \\
Shmelev \& Zychlinski (2025)      & $c$ & \checkmark & -- & \checkmark & --         & $\approx$ \\
\hline
\textbf{This thesis (Model~$A$)}    & \textbf{1} & \checkmark & -- & --  & \checkmark & \checkmark \\
\textbf{This thesis (Model~$B_2$)}& \textbf{1} & \checkmark & \checkmark & -- & \checkmark & \checkmark \\
\textbf{This thesis (Model~$C_2$)}& \textbf{1} & \checkmark & -- & \checkmark & \checkmark & \checkmark \\
\hline
\end{tabular}
\end{table}

The thesis occupies a distinctive position: it provides the \emph{exact} bivariate PGF for a single-server non-preemptive priority queue under both jockeying (Model~$B_2$) and abandonment (Model~$C_2$), combining the queue-length focus of \citet{zuk2023explicit} with the departure mechanisms studied in the approximate and multi-server literature above. The single common theme connecting this work to all entries in Table~\ref{tab:literature} is the two-class non-preemptive priority discipline; the distinguishing contribution is the exact closed-form or integral-form PGF in the single-server setting with one extra mechanism active at a time.

% [Abate et al. (2000): a methodological complement for future work on tail
%  probabilities; not directly cited in the main analysis chapters but useful
%  for asymptotic queue-length results.]
