\section{Preliminaries}\label{sec:prelim}

This section contains some background material referenced throughout the derivations throughout the thesis. By introducing the building blocks in this section, allowing the subsequent the presented arguments to remain focused on their primary intention. Having some known results as preliminaries will help us keep our derivations cleaner.

The central mathematical object in our derivations will be the (P)PGF of the model in question, as well as the idle probability of the system for some corollaries.


\subsection{\texorpdfstring{$M/M/1$}{M/M/1} queue}\label{ssec:mm1}

The $M/M/1$ queue has Poisson arrivals at rate $\lambda$, exponential service
at rate $\mu$, and a single server.  It is stable if and only if $\rho=\lambda/\mu<1$.
Its steady-state queue-length distribution is geometric:
\begin{equation}
    \pi_n = (1-\rho)\,\rho^n, \quad n \ge 0. \label{eq:mm1:steadystate}
\end{equation}
% [Expand: brief derivation via balance equations, or cite adan2002queueing.
%  State mean queue length E[L] = ρ/(1−ρ) (useful for Little's Law applications).]

The \emph{busy period} $B$ begins when a customer arrives to an idle server
and ends when the server next returns to idle.  Its LST and mean are
\begin{align}
    \widetilde{B}(s) &= \frac{(\mu+\lambda+s) - \sqrt{(\mu+\lambda+s)^2 - 4\mu\lambda}}
                              {2\lambda},
    \label{eq:mm1:bp_lst} \\
    \mathbb{E}[B]    &= \frac{1}{\mu(1-\rho)}. \label{eq:mm1:bp_mean}
\end{align}
% [Note: evaluating $\widetilde{B}$ at $s=\lambda_2(1-y)$ recovers $x^*(y)$,
%  the characteristic root of the two-class PGF equation — see §5 (Model~$A$).]

% ─────────────────────────────────────────────────────────────────────────────
\subsection{\texorpdfstring{$M/M/1{+}M$}{M/M/1+M} queue (Erlang-A)}\label{ssec:erlangA}

% [Expand: model definition — M/M/1 with per-customer patience clock Exp(θ).
%  Always positive recurrent for any λ,μ,θ > 0 (contrast with M/M/1).
%  State the steady-state probabilities (product-form via detailed balance).]

The \emph{busy period} $B_C$ of the $M/M/1+M(\lambda_1,\mu,\theta_1)$ queue
is the class-1 subsystem for Model~$C_2$.  Its LST satisfies
\begin{equation}
    \widetilde{B}_C(s) = \cfrac{\mu}{\lambda_1+\mu+s
      -\cfrac{\lambda_1(\mu+\theta_1)}{\lambda_1+\mu+\theta_1+s
      -\cfrac{\lambda_1(\mu+2\theta_1)}{\lambda_1+\mu+2\theta_1+s-\dotsb}}},
    \label{eq:erlangA:bp_lst}
\end{equation}
\textit{Derivation (first-step analysis).}  Let $b_n(s)$ denote the LST of the busy period starting from $n\ge1$ customers (one in service, $n-1$ waiting), so $\widetilde{B}_C(s)=b_1(s)$.  Conditioning on the first event and exploiting exponential memorylessness:
\begin{alignat*}{2}
  n=1\colon\quad
    & (\lambda_1+\mu+s)\,b_1(s) &\;=\;& \mu \;+\; \lambda_1\,b_2(s), \\
  n\ge2\colon\quad
    & (\lambda_1+\mu+(n{-}1)\theta_1+s)\,b_n(s) &\;=\;&
      (\mu+(n{-}1)\theta_1)\,b_{n-1}(s) \;+\; \lambda_1\,b_{n+1}(s).
\end{alignat*}
In the $n=1$ equation there are no waiting customers, so abandonment plays no role; in the $n\ge2$ equation the combined rate of going from $n$ to $n-1$ customers is $\mu+(n-1)\theta_1$ (service completion or any one of the $n-1$ waiting customers abandoning).  Define the ratio $r_n(s)=b_n(s)/b_{n-1}(s)$ for $n\ge2$.  Dividing the $n\ge2$ equation by $b_{n-1}$ gives
\begin{equation}
  r_n(s) \;=\; \cfrac{\mu+(n-1)\theta_1}{\lambda_1+\mu+(n-1)\theta_1+s \;-\; \lambda_1\,r_{n+1}(s)}, \qquad n\ge 2,
  \label{eq:erlangA:r_recursion}
\end{equation}
and the $n=1$ equation rewrites as $b_1(s)=\mu\,/\,(\lambda_1+\mu+s-\lambda_1 r_2(s))$.  The partial denominators $\lambda_1+\mu+(n-1)\theta_1+s$ grow without bound as $n\to\infty$, so the continued fraction converges; substituting the forward recursion~\eqref{eq:erlangA:r_recursion} into $b_1$ and iterating recovers~\eqref{eq:erlangA:bp_lst}.

and its mean admits the closed form
\begin{equation}
    \mathbb{E}[B_C] = \frac{1}{\mu}\,
        {}_1F_1\!\left(1;\,\frac{\mu}{\theta_1}+1;\,\frac{\lambda_1}{\theta_1}\right),
    \label{eq:erlangA:bp_mean}
\end{equation}
where ${}_1F_1$ is defined in \S\ref{ssec:kummer}.
In the limit $\theta_1\to0^+$, $\mathbb{E}[B_C]\to(\mu-\lambda_1)^{-1}$,
recovering \eqref{eq:mm1:bp_mean}.

% [Note: the stability condition for Model~$C$₂,
%  λ₂ · E[B_C] < 1, follows immediately from this mean.]

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Non-preemptive priority and PASTA}\label{ssec:priority}

% [Expand: define the two-class non-preemptive priority discipline.
%  Explain that Class-2 service is never interrupted once started.]

Under Poisson arrivals and a common exponential service rate $\mu$, the
following decomposition holds:
\begin{itemize}
    \item \textbf{Class-1 subsystem.}  Class-1 customers are unaffected by
    Class-2 and form an independent $M/M/1(\lambda_1,\mu)$ queue
    (or $M/M/1+M$ when $\theta_1>0$).
    \item \textbf{Class-2 effective service time.}  From the perspective of
    a Class-2 customer, effective service equals a busy period of the
    Class-1 subsystem.  Class-2 therefore operates as an $M/G/1$ queue with
    service time $B$ (Model~$A$) or $B_C$ (Model~$C_2$) and arrival rate $\lambda_2$.
\end{itemize}

\textbf{PASTA (Poisson Arrivals See Time Averages).}  Because Class-2 arrivals
form a Poisson process, an arriving Class-2 customer sees the stationary
distribution:
\begin{equation}
    \mathbb{P}(\text{arriving Class-2 sees state }s) = \pi(s). \label{eq:prio:pasta}
\end{equation}
This property justifies the conditional-expectation decomposition
\begin{equation}
    P_y(y) = \mathbb{P}(\text{busy},\,N_1=0)\cdot\mathbb{E}[y^{N_2}\mid\text{busy},\,N_1=0]
    \label{eq:prio:py_decomp}
\end{equation}
used in the probabilistic proofs of Models~A and~C$_2$.

% [Note: jockeying breaks the Poisson input seen by each class individually,
%  so PASTA no longer implies the M/G/1 structure — this is why the
%  probabilistic route is unavailable for Models $B$ and B₂.]

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Pollaczek--Khinchine formula}\label{ssec:pk}

% [Expand: brief definition of M/G/1 queue; state stability condition ρ_B < 1.]

For an $M/G/1$ queue with arrival rate $\lambda$, service time $B$ with
LST $\widetilde{B}(s)$, and load $\rho_B = \lambda\,\mathbb{E}[B] < 1$,
the PGF of the stationary number of customers in the system is
\begin{equation}
    L(z) \;=\; \frac{(1-\rho_B)(1-z)\,\widetilde{B}(\lambda(1-z))}
                    {\widetilde{B}(\lambda(1-z)) - z}.
    \label{eq:pk:pgf}
\end{equation}

% [Cite: Kleinrock (1975) or Adan & Resing (2002).]

This formula is applied with two different service-time distributions:
\begin{itemize}
    \item \textbf{Model~$A$}: $B$ = busy period of $M/M/1(\lambda_1,\mu)$,
    so $\widetilde{B}(\lambda_2(1-y)) = x^*(y)$ (see \S\ref{ssec:mm1}).
    \item \textbf{Model~$C_2$}: $B = B_C$ = busy period of
    $M/M/1+M(\lambda_1,\mu,\theta_1)$, so
    $\widetilde{B}_C(\lambda_2(1-y)) = \zeta(y)$ (see \S\ref{ssec:erlangA}).
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Confluent hypergeometric function}\label{ssec:kummer}

The \emph{Kummer confluent hypergeometric function of the first kind} is
\begin{equation}
    {}_1F_1(a;\,b;\,z) \;=\;
        \sum_{n=0}^{\infty} \frac{a^{(n)}}{b^{(n)}\,n!}\,z^n,
    \qquad a^{(n)} = \prod_{k=0}^{n-1}(a+k),\quad a^{(0)}=1,
    \label{eq:kummer:series}
\end{equation}
where $b \notin \{0,-1,-2,\dots\}$ and the series converges for all
$z\in\mathbb{C}$.

% [Expand: mention Kummer's ODE  z u'' + (b-z)u' - a u = 0; cite DLMF §13
%  and Abramowitz & Stegun Ch. 13.]

Two identities used in this thesis:

\textbf{Euler integral representation} ($a,b>0$):
\begin{equation}
    {}_1F_1(a;\,a+b;\,z) \;=\; \frac{1}{B(a,b)}
        \int_0^1 t^{a-1}(1-t)^{b-1} e^{zt}\,dt,
    \label{eq:kummer:euler}
\end{equation}
where $B(a,b)=\Gamma(a)\Gamma(b)/\Gamma(a+b)$.
This is invoked in \S\ref{ssec:C2:anal} to identify the integrals
$\mathcal{J}_0,\mathcal{J}_1$ as Kummer functions evaluated at $s=\lambda_2(1-y)$.

\textbf{Contiguous-function ratio.}  The combination
\begin{equation}
    \frac{{}_1F_1(a+1;\,b;\,z)}{{}_1F_1(a;\,b;\,z)}
    \label{eq:kummer:ratio}
\end{equation}
appears in the closed-form expression for $P_y(y)$ in Model~$B_2$
and in the LST \eqref{eq:erlangA:bp_lst} of the $M/M/1+M$ busy period.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{First-order linear ODEs and PDEs}\label{ssec:ode}

\subsubsection{Integrating factor (variation of parameters)}

% [Expand: one sentence on when this applies — the fundamental PGF equation
%  reduces to an ODE in x for each fixed y when γ₂=θ₂=0 (Models $B$₂ and C₂).]

The standard form of a first-order linear ODE in $x$ (with $y$ as a parameter) is
\begin{equation}
    \frac{dP}{dx} + p(x;\,y)\,P \;=\; q(x;\,y). \label{eq:ode:standard}
\end{equation}
Its integrating factor $\mu_I(x;y) = \exp\!\bigl(\int_0^x p(t;y)\,dt\bigr)$
converts \eqref{eq:ode:standard} to $\frac{d}{dx}[P\cdot\mu_I] = q\cdot\mu_I$,
giving
\begin{equation}
    P(x,y) \;=\; \frac{1}{\mu_I(x;\,y)}
        \left[C(y) + \int_0^x q(t;\,y)\,\mu_I(t;\,y)\,dt\right].
    \label{eq:ode:variation}
\end{equation}
When $\mu_I(0;y)=0$ the constant $C(y)=0$ and the lower boundary term vanishes;
this is the case in both Models~B$_2$ (where $\mu_I \propto x^\alpha(x-y)^\beta$)
and C$_2$ (where $\mu_I \propto x^\alpha(1-x)^{\beta(y)}$).

% [Note: the condition pinning P_y(y) — analyticity at x=y for B₂,
%  boundedness at x=1 for C₂ — is imposed after \eqref{eq:ode:variation}
%  is written down.]

\subsubsection{Method of characteristics}\label{sssec:moc}

% [Expand: one paragraph on the general idea — reduce a first-order PDE to
%  a family of ODEs along characteristic curves in the (x,y)-plane.]

A first-order quasilinear PDE
$a\,P_x + b\,P_y = cP + d$
is solved by tracing \emph{characteristic curves} $(x(t),y(t))$ satisfying
$\dot x = a$, $\dot y = b$.  Along each curve, $P$ satisfies
$\dot P = cP + d$, an ODE in the single variable $t$.

In the two-class PGF equation with jockeying, the convection coefficients
$a = \gamma_1 x(x-y)$ and $b = -\gamma_2 x(x-y)$ give the conserved quantity
\begin{equation}
    C_1 = \gamma_2\,x + \gamma_1\,y \quad (\text{constant along each characteristic}).
    \label{eq:ode:char_constant}
\end{equation}
The general homogeneous solution takes the form
$P_h = F(C_1)\cdot(\text{particular factor})$
where $F$ is an arbitrary function determined by the boundary condition on $P_y(y)$.

% [Note: for Model~$B$ (γ₂>0) determining F requires solving a Volterra integral
%  equation that does not admit a closed form — this is why Model~$B$ is left open.
%  For Model~$B$₂ (γ₂=0) the characteristic direction is purely in x, so the
%  PDE degenerates to the ODE \eqref{eq:ode:standard}, handled by §3.6.1.]
