\section{Variant Model $B_2$ ($\gamma_2=0$)}

\begin{theorem}
    Consider the two-class priority queueing system of Model~B with per-customer jockeying
    confined to the priority class, i.e. $\gamma_1>0$, $\gamma_2=0$ (and
    $\theta_1=\theta_2=0$). Under the stability condition $\rho=\lambda/\mu<1$ with
    $\lambda=\lambda_1+\lambda_2$, the empty-state mass and the boundary generating function
    are given in closed form by
    \begin{equation}
        \pi_0=1-\rho,\qquad \pi(0,0)=\rho(1-\rho),
    \end{equation}
    \begin{equation}
        P_y(y)=\frac{\mu\,\pi(0,0)}{\mu-\lambda y}\cdot
        \frac{{}_1F_1\!\bigl(\alpha(y)+1;\,b^\star(y);\,-\lambda_1 y/\gamma_1\bigr)}
             {{}_1F_1\!\bigl(\alpha(y);\,b^\star(y);\,-\lambda_1 y/\gamma_1\bigr)},
    \end{equation}
    and the bivariate probability generating function, for $0\le x\le y$, by
    \begin{equation}
        P(x,y)=\frac{\mu\,e^{\lambda_1 x/\gamma_1}\,x^{-\alpha}(y-x)^{-\beta}}{\gamma_1 y}
        \left[\frac{\mu\,\pi(0,0)}{\mu-\lambda y}\,
        \frac{{}_1F_1(\alpha+1;b^\star;-\frac{\lambda_1 y}{\gamma_1})}
             {{}_1F_1(\alpha;b^\star;-\frac{\lambda_1 y}{\gamma_1})}\,
        \mathcal J_1(x,y)+(1-y)\pi(0,0)\,\mathcal J_2(x,y)\right],
    \end{equation}
    where ${}_1F_1$ is the confluent hypergeometric (Kummer) function, the exponents are
    \begin{equation}
        \alpha(y)=\frac{\mu}{\gamma_1 y},\qquad
        \beta(y)=\frac{(1-y)(\lambda y-\mu)}{\gamma_1 y},\qquad
        b^\star(y)=\alpha+\beta+1=\frac{\mu+\lambda(1-y)+\gamma_1}{\gamma_1},
    \end{equation}
    and $\mathcal J_1,\mathcal J_2$ are the integrals defined in \eqref{eq:b2_J1}--\eqref{eq:b2_J2};
    on $y<x\le1$ the same expression holds with $\mathcal J_1,\mathcal J_2$ read as their
    finite-part continuations.
\end{theorem}

Let $\gamma_2=0$, then our fundamental equation states:
\begin{equation}
    \bigl[(\lambda_1+\lambda_2+\mu)xy-\mu y-\lambda_1 x^2 y-\lambda_2 xy^2\bigr]P + \gamma_1 xy(x-y)\frac{\partial P}{\partial x} = \mu(x-y)P_y(y) - \mu x(1-y)\pi(0,0).
\end{equation}
Due to the absence of a $\frac{\partial P}{\partial y}$ term, for any fixed $y$, we will have a first-order linear ODE in $x$. Now, for $y\neq0$, we have
\begin{equation}
    \bigl[(\lambda_1+\lambda_2+\mu)x-\mu -\lambda_1 x^2 -\lambda_2 xy\bigr]P + \gamma_1 x(x-y)\frac{\partial P}{\partial x} = \frac{\mu(x-y)P_y(y)}{y} - \frac{\mu x(1-y)\pi(0,0)}{y}.
\end{equation}
Next, we leave the ODE in standard form by dividing both sides by $\gamma_1 x(x-y)$:
\begin{align}
    \begin{split}
        \frac{dP}{dx} + \underbrace{\frac{(\lambda_1+\lambda_2+\mu)x - \mu - \lambda_1 x^2 - \lambda_2 xy}{\gamma_1 x(x-y)}}_{p(x;\,y)}\,P &= \underbrace{\frac{\mu\,P_y(y)}{\gamma_1 xy} - \frac{\mu(1-y)\pi(0,0)}{\gamma_1 y(x-y)}}_{q(x;\,y)} 
    \end{split} \\
    \begin{split}
        \frac{dP}{dx} + p(x;y) P &= q(x;y)
    \end{split}
\end{align}
Here we can write the numerator of $p(x;y)$ as we already know its roots —we write $x^*(y)$ for the taking the negative sign (Equation \eqref{eq:xstar}) and $x_+(y)$ for the root with the positive sign. Hence, we write $(\lambda_1+\lambda_2+\mu)x - \mu - \lambda_1 x^2 - \lambda_2 xy = -\lambda_1 (x-x^*(y))(x-x^+(y))$.
\begin{equation}
    \frac{dP}{dx} - \frac{\lambda_1 (x-x^*(y))(x-x^+(y))}{\gamma_1 x(x-y)}P = q(x;y).
\end{equation}
We can carry on Partial Fraction Decomposition on $p(x;y)$:
\begin{align}
    - \frac{\lambda_1 (x-x^*(y))(x-x^+(y))}{ x(x-y)} = A + \frac{B}{x} +\frac{C}{x-y},
\end{align}
where $A=-\lambda_1$ is obtained by taking the leading coefficient of the $x^2$-term, $B=\frac{\mu}{y}$ by setting $x=0$, and $C=\frac{(1-y)((\lambda_1+\lambda_2)y-\mu)}{y}$ is obtained by setting $x=y$. So
\begin{equation}
    \frac{-\lambda_1(x-x^*)(x-x_+)}{x(x-y)} = -\lambda_1 + \frac{\mu/y}{x} + \frac{(1-y)(\lambda y-\mu)/y}{x-y}.
\end{equation}
If we integrate term by term we obtain
\begin{align}
    \begin{split}
        \int p(x;\,y)\,dx &= \frac{1}{\gamma_1}\!\left[-\lambda_1 x + \frac{\mu}{y}\ln x + \frac{(1-y)(\lambda y-\mu)}{y}\ln(x-y)\right]
    \end{split} \\
    \begin{split}
        \mu_I(x;\,y) &= e^{-\lambda_1 x/\gamma_1}\cdot x^{\,\alpha(y)}\cdot(x-y)^{\beta(y)},
    \end{split}
\end{align}
where
\begin{align}
    \alpha(y)&=\frac{\mu}{\gamma_1 y} \\
    \beta(y)&=\frac{(1-y)(\lambda y-\mu)}{\gamma_1 y}.
\end{align}
Integrating the standard identity $\frac{d}{dx}P\cdot \mu_I= q \cdot \mu_I$ on the interval $(0,x]$, we get
\begin{equation}
    P(x,y)\,\mu_I(x;\,y) - P(0,y)\,\mu_I(0;\,y) = \int_0^x q(t;\,y)\,\mu_I(t;\,y)\,dt.
\end{equation}
Note $\mu_I(x;y) \to 0 \quad$ as $x\to0^+$, so the boundary term vanishes:
\begin{equation}
    P(x,y)\,\mu_I(x;\,y)
= \int_0^x q(t;\,y)\,\mu_I(t;\,y)\,dt.
\end{equation}
Now we substitute $q(t;y)=\frac{\mu\,P_y(y)}{\gamma_1 ty}
  - \frac{\mu(1-y)\pi(0,0)}{\gamma_1 y(t-y)}$ and divide by $\mu_I(x;y)$. Note $P_y(y)$ acts as a constant inside our integral, since it only contains terms on $y$. We can define the integrals:
\begin{align}
    \mathcal{I}_1(x,y) &= \int_0^x e^{-\lambda_1 t/\gamma_1}\,t^{\,\alpha-1}(t-y)^{\beta}\,dt, \\
    \mathcal{I}_2(x,y) &= \int_0^x e^{-\lambda_1 t/\gamma_1}\,t^{\,\alpha}(t-y)^{\beta-1}\,dt
\end{align}
So we obtain
\begin{equation}
    P(x,y) =
\frac{e^{\lambda_1 x/\gamma_1}\,x^{-\alpha(y)}\,(x-y)^{-\beta(y)}}{\gamma_1 y}
\Bigl[\,\mu\,P_y(y)\,\mathcal{I}_1(x,y)
\;-\;\mu(1-y)\pi(0,0)\,\mathcal{I}_2(x,y)\,\Bigr]
\end{equation}

Before extracting $P_y(y)$, it is worth recording that the empty-state mass is available in
closed form, in contrast with Model~C-$2$. Setting $y=x$ in the fundamental equation, the
convection term $\gamma_1 xy(x-y)\partial_x P$ vanishes identically through its factor $(x-y)$,
and so does $\mu(x-y)P_y(y)$; this is the analytic reflection of the fact that jockeying merely
relabels a waiting customer and therefore conserves the total number in the system. What
remains is
\begin{equation}
    \bigl[(\lambda_1+\lambda_2+\mu)x-\mu-(\lambda_1+\lambda_2)x^2\bigr]P(x,x)=-\mu(1-x)\,\pi(0,0),
\end{equation}
whose bracket factors as $(\mu-\lambda x)(x-1)$, so that after cancelling $(x-1)$ we recover the
total-customer generating function of an $M/M/1(\lambda,\mu)$ queue,
\begin{equation}
    P(x,x)=\frac{\mu\,\pi(0,0)}{\mu-\lambda x}=\frac{\pi(0,0)}{1-\rho x}.
\end{equation}
Evaluating at $x=1$ and using $P(1,1)=1-\pi_0=\rho$ with $\pi_0=1-\rho$ yields
\begin{equation}
    \pi(0,0)=\rho(1-\rho),\qquad \pi_0=1-\rho,
    \label{eq:b2_pi00}
\end{equation}
which we shall substitute once $P_y(y)$ has been determined.

The convection coefficient of the ODE is $\gamma_1 x(x-y)$, so $p(x;y)$ has simple poles only
at $x=0$ and at the interior point $x=y\in(0,1)$, and, crucially, not at $x=1$. This is the
structural difference with Model~C-$2$, whose coefficient $\theta_1 x(x-1)$ has its pole at the
boundary $x=1$. Under the stability condition $\rho=\lambda/\mu<1$ we have $\mu-\lambda y>0$ for
every $y\in[0,1]$, hence
\begin{equation}
    \beta(y)=\frac{(1-y)(\lambda y-\mu)}{\gamma_1 y}<0
    \qquad (0<y<1).
    \label{eq:b2_beta_negative}
\end{equation}
Consequently $\mu_I(x;y)^{-1}\propto (x-y)^{-\beta}=(x-y)^{|\beta|}\to0$ as $x\to y$: the
singular homogeneous mode does not blow up, so a boundedness argument à la Model~C-$2$ yields no
condition. The operative requirement is the stronger one that a probability generating function
is holomorphic on the open unit polydisc; we therefore exploit analyticity at $x=y$.

A preliminary technical point concerns the meaning of the integrals on the physical range. For
$0<t<x<y$ the base $t-y$ is negative, so $(t-y)^{\beta}$ is defined only up to a branch. Writing
$(t-y)^{\beta}=e^{i\pi\beta}(y-t)^{\beta}$, the common phase $e^{i\pi\beta}$ cancels in
$P=\mu_I^{-1}\!\int q\,\mu_I$, while the relative phase between the $(t-y)^{\beta}$ and
$(t-y)^{\beta-1}$ contributions is $e^{i\pi\beta}/e^{i\pi(\beta-1)}=-1$, which converts the minus
sign above into a plus. The manifestly real solution on $0<x<y$ is therefore
\begin{equation}
    P(x,y)=\frac{\mu\,e^{\lambda_1 x/\gamma_1}\,x^{-\alpha}(y-x)^{-\beta}}{\gamma_1 y}
    \Bigl[\,P_y(y)\,\mathcal J_1(x,y)+(1-y)\pi(0,0)\,\mathcal J_2(x,y)\,\Bigr],
    \label{eq:b2_Pxy_real}
\end{equation}
\begin{align}
    \mathcal J_1(x,y)&=\int_0^x e^{-\lambda_1 t/\gamma_1}\,t^{\alpha-1}(y-t)^{\beta}\,dt,
    \label{eq:b2_J1}\\
    \mathcal J_2(x,y)&=\int_0^x e^{-\lambda_1 t/\gamma_1}\,t^{\alpha}(y-t)^{\beta-1}\,dt.
    \label{eq:b2_J2}
\end{align}

We now examine the local behaviour at the regular singular point $x=y$. There the
partial-fraction form of $p$ gives a simple pole with residue $C/\gamma_1=\beta$, and the
singular part of the source is governed by the second term of $q$, namely $q(x;y)\sim -D/(x-y)$
with $D=\mu(1-y)\pi(0,0)/(\gamma_1 y)$. The local model equation is thus
\begin{equation}
    P'(x,y)+\frac{\beta}{x-y}\,P(x,y)=\frac{-D}{x-y}+(\text{analytic}).
    \label{eq:b2_local_ode}
\end{equation}
Trying a constant particular solution $P\equiv P_*$ forces $\beta P_*=-D$, i.e.
\begin{equation}
    P(y,y)=P_*=-\frac{D}{\beta}
    =\frac{\mu\,\pi(0,0)}{\mu-\lambda y}
    =\frac{\rho(1-\rho)}{1-\rho y},
    \label{eq:b2_regular_value}
\end{equation}
where the last equality uses $\pi(0,0)=\rho(1-\rho)$ from \eqref{eq:b2_pi00}. The general local
solution is this regular value plus a multiple of the homogeneous mode
$(x-y)^{-\beta}=(x-y)^{|\beta|}$. Because $|\beta|\notin\mathbb Z$ in general, that mode is a
branch term: it is bounded (indeed it vanishes) but not analytic. Holomorphy of $P(\cdot,y)$ at
the interior point $x=y$ therefore forces its coefficient to vanish.

In \eqref{eq:b2_Pxy_real} the prefactor $(y-x)^{-\beta}=(y-x)^{|\beta|}$ multiplies the finite
parts of $\mathcal J_1,\mathcal J_2$ at $x=y$ to produce exactly that branch term; eliminating it
is the single linear condition
\begin{equation}
    P_y(y)\,\mathrm{f.p.}\,\mathcal J_1(y,y)
    +(1-y)\,\pi(0,0)\,\mathrm{f.p.}\,\mathcal J_2(y,y)=0,
    \label{eq:b2_nobranch}
\end{equation}
where $\mathrm{f.p.}$ denotes the Hadamard finite part, that is, the analytic continuation of the
divergent integral in its exponent parameter.

It remains to evaluate these finite parts. Substituting $t=ys$ in
\eqref{eq:b2_J1}--\eqref{eq:b2_J2} and using the Euler representation
$\int_0^1 s^{a-1}(1-s)^{b-1}e^{-w s}\,ds=B(a,b)\,{}_1F_1(a;a+b;-w)$, continued analytically in
$b$ (which is precisely the finite part), gives, with $c=\lambda_1/\gamma_1$,
\begin{align}
    \mathrm{f.p.}\,\mathcal J_1(y,y)&=B(\alpha,\beta+1)\,y^{\alpha+\beta}\,
        {}_1F_1\!\bigl(\alpha;\,\alpha+\beta+1;\,-cy\bigr),\\
    \mathrm{f.p.}\,\mathcal J_2(y,y)&=B(\alpha+1,\beta)\,y^{\alpha+\beta}\,
        {}_1F_1\!\bigl(\alpha+1;\,\alpha+\beta+1;\,-cy\bigr).
\end{align}
The common $y^{\alpha+\beta}$ cancels in \eqref{eq:b2_nobranch}, and the Beta-ratio collapses to
$B(\alpha+1,\beta)/B(\alpha,\beta+1)=\alpha/\beta$. Solving for $P_y(y)$ and using
$(1-y)\,(-\alpha/\beta)=\mu/(\mu-\lambda y)$ together with $\pi(0,0)=\rho(1-\rho)$, we arrive at
the closed form
\begin{empheq}[box=\fbox]{align}
    P_y(y)&=\frac{\mu\,\pi(0,0)}{\mu-\lambda y}\cdot
    \frac{{}_1F_1\!\bigl(\alpha+1;\,b^\star(y);\,-\lambda_1 y/\gamma_1\bigr)}
         {{}_1F_1\!\bigl(\alpha;\,b^\star(y);\,-\lambda_1 y/\gamma_1\bigr)},
    \label{eq:b2_Py_final}\\[2pt]
    \text{with}\quad \alpha(y)&=\frac{\mu}{\gamma_1 y},\qquad
    b^\star(y)=\alpha+\beta+1=\frac{\mu+\lambda(1-y)+\gamma_1}{\gamma_1}. \nonumber
\end{empheq}
The prefactor is the regular value $P(y,y)$ of \eqref{eq:b2_regular_value}, while the Kummer
ratio is the boundary correction. As $y\to0^+$ the ratio tends to $1$ and $P_y(0)=\pi(0,0)$,
consistent with $P_y(0)=\sum_{n_2}\pi(0,n_2)0^{n_2}=\pi(0,0)$.

Substituting \eqref{eq:b2_Py_final} into \eqref{eq:b2_Pxy_real} finally gives, for
$0\le x\le y$,
\begin{empheq}[box=\fbox]{equation}
    P(x,y)=\frac{\mu\,e^{\lambda_1 x/\gamma_1}\,x^{-\alpha}(y-x)^{-\beta}}{\gamma_1 y}
    \left[\frac{\mu\,\pi(0,0)}{\mu-\lambda y}\,
    \frac{{}_1F_1(\alpha+1;b^\star;-\frac{\lambda_1 y}{\gamma_1})}
         {{}_1F_1(\alpha;b^\star;-\frac{\lambda_1 y}{\gamma_1})}\,
    \mathcal J_1(x,y)+(1-y)\pi(0,0)\,\mathcal J_2(x,y)\right].
    \label{eq:b2_Pxy_final}
\end{empheq}
The no-branch condition \eqref{eq:b2_nobranch} guarantees that this extends to a single analytic
function across $x=y$, with diagonal value $P(y,y)=\rho(1-\rho)/(1-\rho y)$; on $y<x\le1$ the same
expression holds with $\mathcal J_1,\mathcal J_2$ read as their finite-part
(incomplete-${}_1F_1$) continuations. Since the constant is fixed in closed form by
\eqref{eq:b2_pi00}, the solution \eqref{eq:b2_Pxy_final} is fully explicit.


