\section{Comparison of Models A, B\texorpdfstring{$_2$}{2}, and C\texorpdfstring{$_2$}{2}}
\label{sec:comparison}

The three solved models represent the three tractable specializations of
Model~X: a pure baseline (A), a jockeying extension (B$_2$), and an
abandonment extension (C$_2$).  Each is obtained from Model~X by zeroing
all but one of the extra mechanisms, and each produces a closed-form or
integral-form expression for $P(x,y)$.  Table~\ref{tab:comparison} collects
the key structural facts; the remainder of this section traces the
consequences.

\begin{table}[H]
\centering
\caption{Structural comparison of Models A, B$_2$, and C$_2$.}
\label{tab:comparison}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{0.26\textwidth} p{0.21\textwidth} p{0.21\textwidth} p{0.21\textwidth}}
\hline
 & \textbf{Model A} & \textbf{Model B$_2$} & \textbf{Model C$_2$} \\
\hline
Extra mechanism
 & none
 & jockeying $1\!\to\!2$ ($\gamma_1>0$)
 & class-1 abandonment ($\theta_1>0$) \\
Stability condition
 & $\rho<1$
 & $\rho<1$
 & $\rho_2\cdot{}_1F_1<1$ \\
Fundamental equation
 & algebraic PGF equation
 & 1st-order PDE in $x,y$
 & 1st-order ODE in $x$ \\
Pinning of $P_y(y)$
 & zero of kernel: $x=x^*(y)$
 & analyticity at $x=y$
 & boundedness at $x=1$ \\
Probabilistic route
 & yes (P-K with $B$)
 & no (jockeying breaks Poisson)
 & yes (P-K with $B_C$) \\
$P_y(y)$
 & rational in $x^*(y)$
 & Kummer ratio \eqref{eq:B2:Py}
 & P-K / Kummer \eqref{eq:C2:Py} \\
$P(x,y)$
 & closed form \eqref{eq:A:fundamental}
 & integral form \eqref{eq:B2:Pxy}
 & integral form \eqref{eq:C2:Pxy} \\
$\pi(0,0)$
 & $\rho(1-\rho)$
 & $\rho(1-\rho)$
 & $(\rho_1{+}\rho_2)\pi_0$ via $\mathbb{E}[B_C]$ \\
$\pi_0$
 & $1-\rho$
 & $1-\rho$
 & $>1-\rho$ (abandonment increases idle time) \\
$P(z,z)$
 & $\rho(1-\rho)/(1-\rho z)$
 & $\rho(1-\rho)/(1-\rho z)$
 & no closed form \\
$\mathbb{E}[N_1]+\mathbb{E}[N_2]$
 & $\rho^2/(1-\rho)$
 & $\rho^2/(1-\rho)$
 & $<\rho^2/(1-\rho)$ \\
$\mathbb{E}[N_1]$ closed form
 & $\rho\rho_1/(1-\rho_1)$
 & no
 & no \\
\hline
\end{tabular}
\end{table}

\subsection*{Model A: the rational baseline}

Model~A imposes no extra departure mechanism.
Class-1 operates as an independent $M/M/1(\lambda_1,\mu)$ queue;
class-2 sees an $M/G/1$ with service time equal to a class-1 busy period
(by PASTA and the priority decomposition of \S\ref{ssec:priority}).
The Pollaczek--Khinchine formula~\eqref{eq:pk:pgf} delivers $P_y(y)$
as a rational function of $x^*(y)$, and $P(x,y)$ is a ratio of polynomials
in $x$, $y$, and $x^*(y)$.  All performance measures --- $\pi_0$, $\pi(0,0)$,
$\mathbb{E}[N_i]$, $\mathbb{E}[W_i]$ --- are in closed form.  Model~A also
furnishes the structural identities that persist across the hierarchy:
$P(z,z)=\rho(1-\rho)/(1-\rho z)$ (total queue is $M/M/1$),
$\pi_0=1-\rho$, and $\pi(0,0)=\rho(1-\rho)$.

\subsection*{Effect of jockeying: Model A $\to$ Model B$_2$}

Adding one-way jockeying ($\gamma_1>0$, $\gamma_2=0$) couples the two queues:
a class-1 customer waiting in Queue~1 may defect to Queue~2 at rate $\gamma_1$.
The consequences are asymmetric.

\emph{What is preserved.}
Jockeying transfers customers but does not remove them, so the total-customer
process remains a standard $M/M/1(\lambda,\mu)$ birth--death chain.
Consequently $P(z,z)$, $\pi_0$, $\pi(0,0)$, and $\mathbb{E}[N]$ are identical
to Model~A for all $\gamma_1>0$.

\emph{What is destroyed.}
Jockeying breaks the per-class Poisson/renewal structure: a class-2 customer
can no longer identify its effective service time with a class-1 busy period,
because class-1 customers now arrive at Queue~2 from inside the system as well
as from outside.  The P-K route is therefore unavailable.

\emph{What is new.}
The fundamental PGF equation gains the convection term
$\gamma_1 xy(x-y)\partial_x P$, which reduces it to a first-order ODE in $x$
for each fixed $y$ (since $\gamma_2=0$ kills the $\partial_y$ term).
The solution closes via an integrating factor with a branch singularity at the
\emph{interior} point $x=y$; the branch coefficient must vanish for $P$ to be
holomorphic, and this analyticity condition pins $P_y(y)$ as a Kummer ratio.
The individual means $\mathbb{E}[N_1]$ and $\mathbb{E}[N_2]$ change with
$\gamma_1$ (class-1 becomes shorter, class-2 longer), with no closed form for
the split.

\subsection*{Effect of abandonment: Model A $\to$ Model C$_2$}

Adding class-1 abandonment ($\theta_1>0$, $\theta_2=0$, $\gamma_i=0$)
allows each waiting class-1 customer to leave the system at rate $\theta_1$.

\emph{What is preserved.}
Class-2 customers still see Poisson arrivals (abandonment only affects
class-1; the class-2 arrival process is undisturbed).
PASTA therefore still applies, and the $M/G/1$ subsystem for class-2 is
intact --- now with effective service time $B_C$ (the busy period of the
$M/M/1+M$ queue of \S\ref{ssec:erlangA}) replacing $B$.
The P-K formula~\eqref{eq:pk:pgf} applies with $B_C$ and delivers $P_y(y)$.

\emph{What is destroyed.}
Abandonment is a true departure: it reduces the total number of customers and
breaks the conservation property.  Consequently $P(z,z)$ no longer has the
simple $M/M/1$ form~\eqref{eq:A:Pzz}, $\pi_0>1-\rho$, $\pi(0,0)$ requires
$\mathbb{E}[B_C]$, and $\mathbb{E}[N]<\rho^2/(1-\rho)$.  The stability
condition is also generalized: $\rho<1$ is no longer necessary; the relevant
condition is $\lambda_2\,\mathbb{E}[B_C]<1$, which is strictly weaker for all
$\theta_1>0$.

\emph{What is new.}
The fundamental equation gains the convection term
$\theta_1 xy(x-1)\partial_x P$, which already reduces it to a first-order ODE
in $x$ for each fixed $y$ (with $\theta_2=0$ killing the $\partial_y$ term).
The integrating factor now has its singularity at the \emph{boundary} $x=1$
with a positive exponent, so $\mu_I$ vanishes there and mere boundedness of
$P$ forces $\int_0^1 q\,\mu_I\,dt=0$.  This determines $P_y(y)$ without any
reference to holomorphy.  Both $\mathbb{E}[N_1]$ and $\mathbb{E}[N_2]$ decrease
relative to Model~A as $\theta_1$ increases: class-2 benefits indirectly
because class-1 abandonments free the server sooner.

\subsection*{Structural comparison of Models B$_2$ and C$_2$}

Both models reduce the fundamental equation to a first-order linear ODE in $x$
and both close in confluent hypergeometric functions; yet the mechanism that
pins $P_y(y)$ is structurally different.  The contrast is as follows.

\begin{center}
\renewcommand{\arraystretch}{1.45}
\begin{tabular}{p{0.28\textwidth} p{0.31\textwidth} p{0.31\textwidth}}
\hline
 & \textbf{Model C$_2$ ($\theta_1>0$)} & \textbf{Model~B$_2$ ($\gamma_1>0$)} \\
\hline
Convection coefficient
 & $\theta_1\,x(x-1)$
 & $\gamma_1\,x(x-y)$ \\
Operative singular point
 & boundary $x=1$
 & interior $x=y$ \\
Exponent at singularity
 & $\beta(y)=\lambda_2(1-y)/\theta_1>0$ \;($\mu_I\to0$)
 & $\beta(y)=(1-y)(\lambda y-\mu)/(\gamma_1 y)<0$ \;($\mu_I\to\infty$) \\
Pinning principle
 & \emph{boundedness}: $\int_0^1 q\,\mu_I\,dt=0$
 & \emph{analyticity}: branch coefficient $=0$ \\
$\pi(0,0)$
 & $M/G/1{+}$renewal (conservation broken)
 & $\rho(1-\rho)$ exactly (conservation holds) \\
Probabilistic backbone
 & class-$2$ is $M/G/1$; P-K with $B_C$
 & class-$1$ is $M/M/1{+}M$ (reneging $=\gamma_1$); no P-K \\
\hline
\end{tabular}
\end{center}

In words: in Model~C$_2$ the abandonment convection $\theta_1 x(x-1)$ places
the singularity at the boundary $x=1$ with a positive integrating-factor
exponent, so $\mu_I$ vanishes there and mere boundedness forces the vanishing
integral that determines $P_y(y)$.  The Poisson character of the class-2
stream survives abandonment, so P-K delivers $P_y(y)$ probabilistically and
an $M/G/1$ renewal argument delivers $\pi(0,0)$.  In Model~B$_2$, the
jockeying convection $\gamma_1 x(x-y)$ places the singularity at the interior
point $x=y$ with a negative exponent, so $\mu_I$ blows up, the homogeneous
mode vanishes, and only the stronger requirement of holomorphy --- the absence
of the non-integer branch $(x-y)^{|\beta|}$ --- pins $P_y(y)$.  Jockeying
destroys the per-class Poisson/renewal structure, so no P-K identity is
available; what remains probabilistic is the total-customer $M/M/1$ behaviour
and the $M/M/1{+}M$ reading of the class-1 marginal, which is the origin of
the ${}_1F_1$ factors.

\subsection*{Common Erlang-A structure}

A common object appears in both B$_2$ and C$_2$:
\begin{equation}
    {}_1F_1\!\left(1;\,\frac{\mu}{r}+1;\,\frac{\lambda_1}{r}\right),
    \label{eq:comp:1F1}
\end{equation}
with $r=\gamma_1$ (Model~B$_2$) or $r=\theta_1$ (Model~C$_2$).
This is the Erlang-A utilisation factor: in C$_2$ it equals $\mu\,\mathbb{E}[B_C]$
(the mean class-1 busy period scaled by $\mu$), while in B$_2$ it controls the
Kummer ratio in $P_y(y)$ through the exponent $\alpha(y)=\mu/(\gamma_1 y)$.
In both cases the function evaluates at a negative argument
$(-\lambda_1/r)$ in C$_2$ and $(-\lambda_1 y/\gamma_1)$ in B$_2$, and the
ratio ${}_1F_1(\alpha+1;\cdot;\cdot)/{}_1F_1(\alpha;\cdot;\cdot)$ acts as a
boundary correction to the Model-A rational formula.

Model~B$_2$ degenerates to Model~A only in the singular limit
$\gamma_1\to0^+$, in which $\alpha,\beta\to\infty$ and the Kummer ratio
collapses to the rational expression $x^*(y)(1-y)/(x^*(y)-y)$ of Model~A;
this limit is not uniform, consistent with $\gamma_1$ multiplying the
highest-order term of the ODE.  Model~C$_2$ degenerates to Model~A uniformly
as $\theta_1\to0^+$: $\mathbb{E}[B_C]\to(\mu-\lambda_1)^{-1}$, and all
C$_2$ formulas recover their Model-A counterparts.
